Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-07 arxiv/proof_examples.tex
revise

\paragraph{Part $1$: the reduction.}
Under the two-task setting of Corollary \ref{cor_MTL_loss}, the optimization objective \eqref{eq_mtl} becomes
		f(A_1,A_2, B) =   \bignorm{X^{(1)} B A_1 - Y^{(1)}}^2+ \bignorm{X^{(2)} B A_2 - Y^{(2)}}^2, \label{eq_mtl_2task_cov}
	where $B \in \real^{p}$ and $A_1, A_2\in \R$ because $r=1$ in the two-task case. % for $1 < r < t$ by our assumption.
	Without loss of generality, we assume that both $A_1$ and $A_2$ are nonzero, since otherwise we can add a small perturbation $\delta$ to them and later remove this perturbation by taking $\delta\to 0$ in the final result. Again using the local optimality condition ${\partial f}/{\partial B} = 0$, for any $A_1$ and $A_2$ we obtain the optimal $\hat{B}$ as follows
		\hat{B} \define  (A_1^2 (X^{(1)})^{\top}X^{(1)} + A_2^2 (X^{(2)})^{\top}X^{(2)})^{-1} (A_1 (X^{(1))})^{\top}Y^{(1)} + A_2 (X^{(2)})^{\top}Y^{(2)}). \label{eq_Bhat_2task} %\\
		%&= (B^\star A ^{\top}) (A A^{\top})^{-1} + (X^{\top}X)^{-1}X^{\top}   \bigbrace{\sum_{j=1}^t \varepsilon_i A_i^{\top}} (A  A^{\top})^{-1}.
 Now inserting $\hat B$ back into equation \eqref{eq_mtl_2task_cov}, we obtain an objective that only depends on $x:=A_1/A_2$ as follows %\HZ{$A$ has been used to denote the output layers. Could you replace $A$ with another symbol (say $x$)?}
		 g(x) \define & \left\| X^{(1)} \Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (x\beta_2-\beta_1)+ \left(x^2 X^{(1)}\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1}\right)\epsilon^{(1)} \right. \nonumber\\
		& \left. + x X^{(1)}\Sigma(A)^{-1} (X^{(2)})^\top \epsilon^{(2)} \right\|^2 \nonumber\\
		 + & \left\| X^{(2)} \Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} (x\beta_1-x^2\beta_2)+ \left(X^{(2)}\Sigma(x)^{-1} (X^{(2)})^\top-\id_{n_2}\right)\epsilon^{(2)} \right. \nonumber\\
		 &\left. + x X^{(2)}\Sigma(x)^{-1} (X^{(1)})^\top \epsilon^{(1)} \right\|^2, 
		\label{eq_mtl_A12}
	where we denote {\cor revise}
	$$ \Sigma(x):= x^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$$
We can calculate the partial expectation of $g(A)$ over the randomness in $\epsilon_1$ and $\epsilon_2$:
\HZ{could you align the Eq so that they look better on page?}
&\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(A) \mid X_1, X_2 ,\beta_1,\beta_2} \\
= &\left\| X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)\right\|^2 +\left\| X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)\right\|^2  \\
& + \exarg{\epsilon^{(1)}}{(\epsilon^{(1)})^\top \left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\epsilon^{(1)}} \\
&+\exarg{ \epsilon^{(2)}}{(\epsilon^{(2)})^\top \left(X^{(2)}\Sigma(A)^{-1} (X^{(2)})^\top-\id_{n_2}\right)^2\epsilon^{(2)}}   \\
&+  A^2 \exarg{ \epsilon^{(2)}}{ (\epsilon^{(2)} )^\top X^{(2)} \Sigma(A)^{-1}  (X^{(1)})^\top X^{(1)}\Sigma(A)^{-1} (X^{(2)})^\top \epsilon^{(2)}} \\
& +A^2 \exarg{ \epsilon^{(1)}} { (\epsilon^{(1)})^\top X^{(1)} \Sigma(A)^{-1}  (X^{(2)})^\top X^{(2)}\Sigma(A)^{-1} (X^{(1)})^\top \epsilon^{(1)}  }\\
		&= \left\| X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)\right\|^2 +\left\| X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)\right\|^2  \\
		&+ \sigma^2\tr\left[ \left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\right] +\sigma^2\tr\left[ \left(X^{(2)}\Sigma(A)^{-1} (X^{(2)})^\top-\id_{n_2}\right)^2 \right]\\
&+ 2 A^2 \sigma^2 \tr\left[ \Sigma(A)^{-1}  (X^{(1)})^\top X^{(1)}\Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} \right]  \nonumber\\
		& = \left\| X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-A\beta_2)\right\|^2 +A^2\left\| X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)} (\beta_1-A\beta_2)\right\|^2 +\sigma^2(n_1+n_2-p).
For the random-effects model, we can further take the partial expectation over $\beta_1$ and $\beta_2$: 
&\exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(A) \mid X_1, X_2} \nonumber\\
&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(2)})^\top X^{(2)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)}\right] \nonumber\\
&+ A^2[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)}\right] +\sigma^2(n_1+n_2-p) \nonumber\\
&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)}\right]+\sigma^2(n_1+n_2-p), \label{EgA}
where in the first step we used that the entries of $\beta_1-A\beta_2\in \R^p$ are i.i.d. Gaussian random variables with mean zero and variance $p^{-1}[(A-1)^2\kappa^2 + (A^2+1)d^2/2]$, and in the second step we used  %Using $(X^{(2)})^\top X^{(2)}=\hat\Sigma- A^2 (X^{(1)})^\top X^{(1)}$, we can further simplify that
 & (X^{(2)})^\top X^{(2)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} + A^2   (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)}  \\
 &=  (X^{(2)})^\top X^{(2)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} + A^2  (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)}  \\
 & =   (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} .
Here in the first step of the above derivation, we used that 
 $$(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} =\left( A^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} = (X^{(2)})^\top X^{(2)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)}.$$

>>>>

\paragraph{Part $1$: the reduction.}
Under the two-task setting of Corollary \ref{cor_MTL_loss}, the optimization objective \eqref{eq_mtl} becomes
	\begin{align}
		f(A_1,A_2, B) =   \bignorm{X^{(1)} B A_1 - Y^{(1)}}^2+ \bignorm{X^{(2)} B A_2 - Y^{(2)}}^2, \label{eq_mtl_2task_cov}
	\end{align}
	where $B \in \real^{p}$ and $A_1, A_2\in \R$ because $r=1$ in the two-task case. % for $1 < r < t$ by our assumption.
	Without loss of generality, we assume that both $A_1$ and $A_2$ are nonzero, since otherwise we can add a small perturbation $\delta$ to them and later remove this perturbation by taking $\delta\to 0$ in the final result. Again using the local optimality condition ${\partial f}/{\partial B} = 0$, for any $A_1$ and $A_2$ we obtain the optimal $\hat{B}$ as follows
	\begin{align}
		\hat{B} \define  (A_1^2 (X^{(1)})^{\top}X^{(1)} + A_2^2 (X^{(2)})^{\top}X^{(2)})^{-1} (A_1 (X^{(1))})^{\top}Y^{(1)} + A_2 (X^{(2)})^{\top}Y^{(2)}). \label{eq_Bhat_2task} %\\
		%&= (B^\star A ^{\top}) (A A^{\top})^{-1} + (X^{\top}X)^{-1}X^{\top}   \bigbrace{\sum_{j=1}^t \varepsilon_i A_i^{\top}} (A  A^{\top})^{-1}.
	\end{align}
 Now inserting $\hat B$ back into equation \eqref{eq_mtl_2task_cov}, we obtain an objective that only depends on $x:=A_1/A_2$ as follows %\HZ{$A$ has been used to denote the output layers. Could you replace $A$ with another symbol (say $x$)?}
 \begin{align}
		 g(x) \define & \left\| X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (x\beta_2-\beta_1)+ \left(x^2 X^{(1)}\hat \Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1}\right)\epsilon^{(1)} \right. \nonumber\\
		& \left. + x X^{(1)}\hat \Sigma(x)^{-1} (X^{(2)})^\top \epsilon^{(2)} \right\|^2 \nonumber\\
		 + & \left\| X^{(2)} \hat \Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} (x\beta_1-x^2\beta_2)+ \left(X^{(2)}\hat\Sigma(x)^{-1} (X^{(2)})^\top-\id_{n_2}\right)\epsilon^{(2)} \right. \nonumber\\
		 &\left. + x X^{(2)}\hat \Sigma(x)^{-1} (X^{(1)})^\top \epsilon^{(1)} \right\|^2, 
		\label{eq_mtl_A12}
	\end{align}
	where we denote $ \hat \Sigma(x):= x^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$ We can calculate the partial expectation of $g(x)$ over the randomness in $\epsilon_1$ and $\epsilon_2$:
%\HZ{could you align the Eq so that they look better on page?}
\begin{align}	
&\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(x) \mid X_1, X_2 ,\beta_1,\beta_2} \nonumber\\
&= \left\| X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-x\beta_2)\right\|^2 \nonumber\\
&+x^2 \left\| X^{(2)} \hat\Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} (\beta_1-x\beta_2)\right\|^2  \nonumber\\
& + \exarg{\epsilon^{(1)}}{(\epsilon^{(1)})^\top \left(x^2 X^{(1)}\hat \Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\epsilon^{(1)}} \nonumber\\
&+\exarg{ \epsilon^{(2)}}{(\epsilon^{(2)})^\top \left(X^{(2)}\hat \Sigma(x)^{-1} (X^{(2)})^\top-\id_{n_2}\right)^2\epsilon^{(2)}}   \nonumber\\
&+  x^2 \exarg{ \epsilon^{(2)}}{ (\epsilon^{(2)} )^\top X^{(2)}\hat \Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)}\hat\Sigma(x)^{-1} (X^{(2)})^\top \epsilon^{(2)}} \nonumber\\
& +x^2 \exarg{ \epsilon^{(1)}} { (\epsilon^{(1)})^\top X^{(1)}\hat \Sigma(x)^{-1}  (X^{(2)})^\top X^{(2)}\hat\Sigma(x)^{-1} (X^{(1)})^\top \epsilon^{(1)}  }. \label{Eg(x)}
\end{align}
Using the following identity
\begin{align*}
(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} & =\left( A^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} \\
&= (X^{(2)})^\top X^{(2)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)},
\end{align*}
we can simplify that 
\begin{align*}
&  \left\| X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-x\beta_2)\right\|^2  +x^2 \left\| X^{(2)} \hat\Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} (\beta_1-x\beta_2)\right\|^2 \\
&=
\end{align*}

Hence \eqref{Eg(x)} can be simplified to 
\begin{align*}
		&= \left\| X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)\right\|^2 +\left\| X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)\right\|^2  \\
		&+ \sigma^2\tr\left[ \left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\right] +\sigma^2\tr\left[ \left(X^{(2)}\Sigma(A)^{-1} (X^{(2)})^\top-\id_{n_2}\right)^2 \right]\\
&+ 2 A^2 \sigma^2 \tr\left[ \Sigma(A)^{-1}  (X^{(1)})^\top X^{(1)}\Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} \right]  \nonumber\\
		& = \left\| X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-A\beta_2)\right\|^2 +A^2\left\| X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)} (\beta_1-A\beta_2)\right\|^2 +\sigma^2(n_1+n_2-p).
\end{align*}
For the random-effects model, we can further take the partial expectation over $\beta_1$ and $\beta_2$: 
\begin{align}	
&\exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(A) \mid X_1, X_2} \nonumber\\
&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(2)})^\top X^{(2)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)}\right] \nonumber\\
&+ A^2[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)}\right] +\sigma^2(n_1+n_2-p) \nonumber\\
&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)}\right]+\sigma^2(n_1+n_2-p), \label{EgA}
\end{align}
where in the first step we used that the entries of $\beta_1-A\beta_2\in \R^p$ are i.i.d. Gaussian random variables with mean zero and variance $p^{-1}[(A-1)^2\kappa^2 + (A^2+1)d^2/2]$, and in the second step we used  %Using $(X^{(2)})^\top X^{(2)}=\hat\Sigma- A^2 (X^{(1)})^\top X^{(1)}$, we can further simplify that
\begin{align*}
 & (X^{(2)})^\top X^{(2)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} + A^2   (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} \Sigma(A)^{-1} (X^{(1)})^\top X^{(1)}  \\
 &=  (X^{(2)})^\top X^{(2)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} + A^2  (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)}  \\
 & =   (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} .
\end{align*}
